---
title: "Week 4 Group Project Using Queens Data"
author: "Group2 - Tony Garza"
date: "June 8, 2016"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown Analysis Report

This is the code used to both scrub and analyze the data. This R Markdown document will culminate with a series of plots using the Queens data. The plots will be used to determine any potential patterns as well as whether or not we are working with a normal distribution at all.

##Load the Data
```{r echo=TRUE}
require(plyr)
require(gdata)
# Read the raw data from rollingsales_queens.xls into variable queens
## Need to download the perl interpreter for Windows (ran this one on a Mac though).  http://strawberryperl.com/
queens <- read.xls("rollingsales_queens.xls", pattern = "BOROUGH")
```
##Explore the Data
Let's view the data to make sure it loaded correctly.
```{r}
head(queens)
```
Let's continue to explore the data:
```{r}
str(queens)
summary(queens)
```
##Clean the Data
Now let's start cleaning up the data: <br/>
1. First let's convert all labels to lower case to make things easier/consistent.
```{r}
names(queens) <- tolower(names(queens))
```
2. Let's make sure our data is all numeric
```{r}
queens$sale.price.n <- as.numeric(gsub("[^[:digit:]]","", queens$sale.price))
count(is.na(queens$sale.price.n))
queens$gross.sqft <- as.numeric(gsub("[^[:digit:]]","",queens$gross.square.feet))
count(is.na(queens$gross.square.feet))
```
3. Let's now keep only the actual sales:
```{r}
queens.sale <- queens[queens$sale.price.n!= 0,]
summary(queens.sale)
```
4. This is still a lot of data points so let's also just look at 1-, 2-, and  3-family homes:
```{r}
queens.homes <- queens.sale[which(grepl("FAMILY", queens.sale$building.class.category)),]
summary(queens.homes)
```
## Analyze and Plot the Data
First let's see what our data looks like now and see if we should scrub the data more:
```{r}
qqplot(log10(queens.homes$gross.square.feet),log10(queens.homes$sale.price.n))
```
<br/>
Looks like we still have quite a few outliers so let's clean the data further and re-plot:
```{r}
queens.homes$outliers <- log10(queens.homes$sale.price.n) <=5 +0
queens.homes <- queens.homes[which(queens.homes$outliers == 0),]
qqplot(log10(queens.homes$gross.square.feet),log10(queens.homes$sale.price.n))
```
<br/>
This piece of the data looks to be close to a normal distribution. Let's try a regular plot to see where the data is tending.
```{r}
plot(log10(queens.homes$gross.square.feet),log10(queens.homes$sale.price.n))
```
<br/>
Let's look at the distribution of the individual data sets of gross sqft and sale price:
```{r}
boxplot(log10(queens.homes$gross.square.feet),log10(queens.homes$sale.price.n))
```
<br/>
Since we are receiving a warning about outliers in box plot one we may want to scrub the data further:
```{r}
queens.homes$outliers2 <- log10(queens.homes$gross.square.feet) <=2.8 +0
queens.homes <- queens.homes[which(queens.homes$outliers2 == 0),]
qqplot(log10(queens.homes$gross.square.feet),log10(queens.homes$sale.price.n))
boxplot(log10(queens.homes$gross.square.feet),log10(queens.homes$sale.price.n))
```
<br/>
No warnings this time. Here is some supporting data for the boxplots:
```{r}
summary(log10(queens.homes$gross.square.feet))
summary(log10(queens.homes$sale.price.n))
```
<br/>
Without the log10 correction you can see this data is naturally skewed to the left:
```{r}
hist(queens.homes$gross.square.feet)
hist(queens.homes$sale.price.n)
```
